---
title: "Homework 1"
author: "Ekaterina Hofrenning"
date: '2023-09-14'
output: github_document
---

Load Libraries:

```{r, message = F}
library(tidyverse)
library(moderndive)
```

# Question 1:
*Write a short description of the dataset using inline R code; accessing the dataset help file can be informative. In your discussion, please include:*

- *the variables in this dataset, including names / values of important variables*
- *the size of the dataset (using nrow and ncol)*
- *the mean temperature*

*Make a scatterplot of temp (y) vs time_hour (x); color points using the humid variable (adding color = ... inside of aes in your ggplot code should help). Describe patterns that are apparent in this plot. Export your scatterplot to your project directory using ggsave.*

```{r}
data("early_january_weather") # load data

names(early_january_weather) # var names
nrow(early_january_weather) # number of rows
ncol(early_january_weather) # number of cols
mean(early_january_weather$temp) # mean temp
```

```{r}
ggplot(early_january_weather, aes(x = temp, y = time_hour, color = humid)) + geom_point()
ggsave("temp_time_humid.png")
```

The `early_january_weather` dataset contains variables about temperature, dew point, humidity, wind direction, wind speed, and wind gust, precipitation, sea level pressure, and visibility metrics in addition to the date and time each was measured in January. In total, there are 358 observations and 15 variables. The mean temperature is 39.58 degrees. There appears to be a slight positive trend between temperature and the date. On each day, the temperature increases and decreases as the hours go on. Some days are more humid than others-- January 12-14th is a stretch of days with high humidity.

# Question 2:

*Create a data frame comprised of:*

- *a random sample of size 10 from a standard Normal distribution*
- *a logical vector indicating whether elements of the sample are greater than 0*
- *a character vector of length 10*
- *a factor vector of length 10, with 3 different factor “levels”*

*Try to take the mean of each variable in your dataframe. What works and what doesn’t?*

```{r, warning=FALSE}
my_data <- # create dataset
  tibble(
  samp = rnorm(10, mean = 0, sd = 1),
  big_0 = samp > 0,
  char = c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J"),
  fact = factor(c("low", "low", "med", "med", "high", "high", "low", "high", "med", "low"))
  )

my_data %>% summarize_all(mean) # calculate means
```

When taking the mean of each variable in my dataframe, the means for the random sample and logical vectors works while the means for the character and factor vectors do not work.

*In some cases, you can explicitly convert variables from one type to another. Write a code chunk that applies the as.numeric function to the logical, character, and factor variables (please show this chunk but not the output). What happens, and why? Does this help explain what happens when you try to take the mean?*

```{r, eval = FALSE, warning = FALSE}
my_data %>% 
  mutate(as.numeric(big_0),
         as.numeric(char),
         as.numeric(fact))
```

After applying as.numeric() to the logical, character, and factor variables, the logical vector turns into 0s and 1s (0=F, 1=T), the character vector fails to convert and becomes NAs, and the factor variable turns into 1s,2s, and 3s. This makes sense because: logical variables are the same if you think about them in terms of 0/1's or T/F's, character vectors are a composition of letters so they can't be thought of in the numeric sense, and factor variables are groups so they could be assigned a number for each group. This helps when thinking about taking the means because character and factor variables don't have ordinal and numeric properties, while the logical vector can be read by the computer as 0s and 1s to take the mean.
