---
title: "p8105_hw1_eih2108"
author: "Ekaterina Hofrenning"
date: '2023-09-14'
output: github_document
---

Load Libraries:

```{r}
library(tidyverse)
library(moderndive)
```

# Question 1:
Write a short description of the dataset using inline R code; accessing the dataset help file can be informative. In your discussion, please include:

- the variables in this dataset, including names / values of important variables
- the size of the dataset (using nrow and ncol)
- the mean temperature

Make a scatterplot of temp (y) vs time_hour (x); color points using the humid variable (adding color = ... inside of aes in your ggplot code should help). Describe patterns that are apparent in this plot. Export your scatterplot to your project directory using ggsave.

```{r}
data("early_january_weather")

names(early_january_weather) # var names
nrow(early_january_weather) # number of rows
ncol(early_january_weather) # number of cols
mean(early_january_weather$temp) # mean temp
```

```{r}
ggplot(early_january_weather, aes(x = temp, y = time_hour, color = humid)) + geom_point()
ggsave("temp_time_humid.png")
```

The `early_january_weather` dataset contains variables about temperature, dew point, humidity, wind direction, wind speed, and wind gust, precipitation, sea level pressure, and visibility metrics in addition to the date and time each was measured in January. In total, there are 358 observations and 15 variables. The mean temperature is 39.58 degrees. There appears to be a slight positive trend between temperature and the date. On each day, the temperature increases and decreases as the hours go on. Some days are more humid than others-- January 12-14th is a stretch of days with high humidity.

# Question 2:



